Variable Density Tesseroids: Gravity fields calculation in spherical coordinates using variable densities in depth

### 2017-11-02 manuscript/manuscript.tex
Added potential and gz in delta determination

\begin{summary}
We present a new open-soruce software that allows to compute the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e. a forward gravity model in spherical coordinates for varying densities on the radial direction.
It is written in Python and Cython language and is compatible with the classes and functions of the library for geophysical modelling Fatiando a Terra.
It numerically approximates the gravitational fields by applying the Gauss-Legendre Quadrature along with a preexisting adaptive discretization algorithm and a new density-based discretization algorithm that control both the accuracy and the computation time of the forward model.
The first one is controlled by a predefined distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
While the second one, regulated by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
We have also obtained analytical solutions for a spherical shell with variable density and compared them with the results of the numerical model in case of a linear, an exponential and a discontinuous density function.
These comparisons allowed to obtain default values for the distance-size and the delta ratio in order to guarantee the accuracy of the forward model.
The resulting optimal values of $D$ for the gravity potential, the components of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
While a $\delta=0.2$ is needed even for high varying exponential densities and discontinuous functions.
We have finally tested the new software by modelling the Neuquen sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning a linear density variation to it and computing all the gravity fields it generates.
\end{summary}

@@

To do this we applied a similar methodology used in the determination of the distance-size ratio $D$: we proposed a set of exponential density functions (as the one in equation \ref{eq:density-exp}) with different values of $b$, and computed the difference between the gravity fields generated by the numerical model with the corresponding analytical solution, each time with a different value of $\delta$ ranging from 0.1 to 1 with a step of 0.1.
This tests have been performed for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
\caption{
    Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
    (a) Density functions used to perform the test.
    (b) Differences between the numerical model and the corresponding analytical solution for several values of $\delta$ and $b$.}
\label{fig:exponential-delta-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
\caption{
    Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
    (a) Density functions used to perform the test.
    (b) Differences between the numerical model and the corresponding analytical solution for several values of $\delta$ and $b$.}
\label{fig:exponential-delta-thick}
\end{figure}

On Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a we can see the exponential density functions used on each test for a thin and thick spherical shell, respectively.
While Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b show their corresponding results.

@@

We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from low to high varying functions.
Each comparison was made for different values of $\delta$, keeping the default values of $D$ previously obtained.
This tests have shown that a value of $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while the same results can be obtained performing this tests with the discontinuous density function.

>>>>

\begin{summary}
We present a new open-soruce software that allows to compute the gravitational fields generated by any tesseroid whose density varies arbitrarily in depth, i.e. a forward gravity model in spherical coordinates for varying densities on the radial direction.
It is written in Python and Cython language and is compatible with the classes and functions of the library for geophysical modelling Fatiando a Terra.
It numerically approximates the gravitational fields by applying the Gauss-Legendre Quadrature along with a preexisting adaptive discretization algorithm and a new density-based discretization algorithm that control both the accuracy and the computation time of the forward model.
The first one is controlled by a predefined distance-size ratio ($D$) and reduces the errors due to the distance between the tesseroid and the computation point.
While the second one, regulated by a delta ratio ($\delta$), decreases the errors introduced by the variation of the density function.
We have also obtained analytical solutions for a spherical shell with variable density and compared them with the results of the numerical model in case of a linear, an exponential and a discontinuous density function.
These comparisons allowed to obtain default values for the distance-size and the delta ratio in order to guarantee the accuracy of the forward model.
The resulting optimal values of $D$ for the gravity potential, the components of its gradient and the Marussi tensor components are 1, 2 and 8, respectively.
While a $\delta=0.2$ is needed even for high varying exponential densities and discontinuous functions.
We have finally tested the new software by modelling the Neuqu\'en sedimentary basin, located to the east of the Andes between 32$^\circ$S and 40$^\circ$S latitude, assigning a linear density variation to it and computing all the gravity fields it generates.
\end{summary}

@@

To do this we applied a similar methodology used in the determination of the distance-size ratio $D$.
Using different values of $\delta$, ranging from 0.1 to 1 with a step of 0.1, we computed the potential, the $g_z$ component of the gradient and the $g_{zz}$ component of the Marussi tensor generated by a spherical shell with several exponential densities (like the one in equation \ref{eq:density-exp}): each one with a different value of $b$.
We performed this tests for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.

Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a show the exponential density functions used on each test for a thin and thick spherical shell, respectively.
While on Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b we can see the differences between the resulting $g_{zz}$ with the corresponding analytical solution.
The results for the other gravity fields were excluded from the figures due to their similarities with the ones for $g_{zz}$.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
\caption{
    Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
    (a) Density functions used to perform the test.
    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
\label{fig:exponential-delta-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
\caption{
    Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
    (a) Density functions used to perform the test.
    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
\label{fig:exponential-delta-thick}
\end{figure}

@@

We have performed similar tests for the density-based discretization algorithm comparing the numerical model with the analytical solutions for a spherical shell with various exponential densities: from low to high varying functions.
Each comparison was made for different values of $\delta$, keeping the default values of $D$ previously obtained.
This tests have shown that a value of $\delta = 0.2$ guarantees an accuracy of 0.1\% or better on every case, while the same results can be obtained performing this tests with the discontinuous density function.
The default value $\delta$ is the same for every gravity field computation.
